---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


```{r ReadData}

  activity<-read.table("activity.csv", header = TRUE, sep = ",")
  
```


## What is mean total number of steps taken per day?

```{r TotalNumberSteps}
  steps.daySum <- aggregate(activity$steps, by=list(date=activity$date), FUN=sum, na.rm = TRUE)
  names(steps.daySum)[names(steps.daySum) == "x"] = "steps"
  library(ggplot2)
  g = ggplot(steps.daySum, aes(x = date, y = steps))+ geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(g)
  steps.dayMean <- aggregate(activity$steps, by=list(date=activity$date), FUN=mean, na.rm = TRUE)
  names(steps.dayMean)[names(steps.dayMean) == "x"] = "steps.daymean"
  
  steps.dayMedian <- aggregate(activity$steps, by=list(date=activity$date), FUN=median, na.rm = TRUE)
  names(steps.dayMedian)[names(steps.dayMedian) == "x"] = "steps.daymedian"

```
Bellow we see the mean per day.
```{r}
  steps.dayMean
```

Bellow we see the median per day.
```{r}
  steps.dayMedian
```

## What is the average daily activity pattern?

```{r IntervalPlots}
  steps.interval <- aggregate(activity$steps, by=list(interval=activity$interval), FUN=sum, na.rm = TRUE)
  names(steps.interval)[names(steps.interval) == "x"] = "steps"

  library(ggplot2)
  g = ggplot(steps.interval, aes(x = interval, y = steps))+ geom_line()
  print(g)
  maxIntervalSteps <- max(steps.interval$steps, na.rm = TRUE)
  maxInterval <- which(steps.interval$steps == max(steps.interval$steps), arr.ind = TRUE)
```

The inerval with the most steps recorded is:
```{r}  
  steps.interval$interval[maxInterval]
```

Bellow we see the steps of all the 5 minute intervals.
```{r}
  steps.interval
```

## Inputing missing values
  
The number of missing values is:
```{r}
  sum(is.na(activity$steps))
  activity.full <- activity
  for (i in 1:nrow(activity.full)){
    if (is.na(activity.full$steps[i])){
      temp.int <- which(steps.interval$interval == activity.full$interval[i])
      activity.full$steps[i] <- steps.interval$steps[temp.int]/61
    }
  }
```
  
The number of missing values now is:
```{r}
  sum(is.na(activity.full$steps))
```

```{r}
  steps.full.daySum <- aggregate(activity.full$steps, by=list(date=activity.full$date), FUN=sum, na.rm = TRUE)
  names(steps.full.daySum)[names(steps.full.daySum) == "x"] = "steps"
  g = ggplot(steps.full.daySum, aes(x = date, y = steps))+ geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(g)
```
After adding the missing values the mean went up by about 15% and the median has some values where before it had everywhere zeros.

```{r}
  steps.full.dayMean <- aggregate(activity.full$steps, by=list(date=activity.full$date), FUN=mean, na.rm = TRUE)
  names(steps.full.dayMean)[names(steps.full.dayMean) == "x"] = "steps.full.daymean"
  
  steps.full.dayMedian <- aggregate(activity.full$steps, by=list(date=activity.full$date), FUN=median, na.rm = TRUE)
  names(steps.full.dayMedian)[names(steps.full.dayMedian) == "x"] = "steps.full.daymedian"
  
```
  
The number of missing values is:
```{r}
  sum(is.na(activity.full$steps))
```

## Are there differences in activity patterns between weekdays and weekends?

```{r Weekdays}

  activity$day <- weekdays(as.Date(activity$date, format='%Y-%m-%d'))
  for (i in 1:nrow(activity)){
    if (activity$day[i] == "Κυριακή" | activity$day[i] == "Σάββατο"){
      activity$day[i] = "weekend"
    }
    else{
      activity$day[i] = "weekday"
    }
  }
  
  steps.interval.day <- aggregate(activity$steps, by=list(interval=activity$interval, day=activity$day), FUN=sum, na.rm = TRUE)
  names(steps.interval.day)[names(steps.interval.day) == "x"] = "steps"
  
  ggplot(steps.interval.day, aes(x = interval, y = steps, group = day)) +
    geom_line() +
    facet_grid( ~ day)
  
```